Mining the human microbiome for clinical insight

### 2019-01-25 24hr/24hr_ms.tex
update 24hr with latest figures

However, an alternative explanation for these spikes in conjugated bile acids may be that they reflect only incidental increases in the sampling of individuals with diarrhea at these times. Diarrhea is characterized by a decrease in the conversion of conjugated primary bile acids to their unconjugated forms. Furthermore, because our metabolomics data is generated from the liquid phase of wastewater after removing solids, diarrhea will contribute relatively more than solid stools to the sampled metabolites at any given time. Thus, our results are consistent with a sampling enriched in individuals with diarrhea at times when they defecate. By contrast, the contribution of stools from healthy individuals (as indicated by the levels of cholate and deoxcholate) remains relatively constant over the course of the day. Taken together, the findings that fecal and urinary metabolites differ in their dynamics and that individual people may disproportionately contribute to individual samples, suggest that developing methods to correct for biases in sampling and normalize for the number of people contributing to a given sample will be required before these data can be acted upon in public health contexts.

@@

\begin{figure}[h]
\begin{center}
    \includegraphics[width=\textwidth]{{24hr/figures/fig5_mining_big_data.png}}
    \caption{Mining untargeted data suggests identification of human health, behavioral, and lifestyle biomarkers. Heatmap of metabolite feature abundance over 24-hour samplings, with putatively identified metabolites labeled. Metabolites labeled with opaque circles were confirmed with analytical standards or MS2 matching. Labels with transparent circles were identified via mass-matching neutral mass with HMDB metabolites.}\label{24hr:fig5}
\end{center}
\end{figure}

>>>>

However, an alternative explanation for these spikes in conjugated bile acids may be that they reflect only incidental increases in the sampling of individuals with diarrhea at these times. Diarrhea is characterized by a decrease in the conversion of conjugated primary bile acids to their unconjugated forms. Furthermore, because our metabolomics data is generated from the liquid phase of wastewater after removing solids, diarrhea will contribute relatively more than solid stools to the sampled metabolites at any given time. Thus, our results are consistent with a sampling enriched in individuals with diarrhea at times when they defecate. By contrast, the contribution of stools from healthy individuals (as indicated by the levels of cholate and deoxycholate) remains relatively constant over the course of the day. Taken together, the findings that fecal and urinary metabolites differ in their dynamics and that individual people may disproportionately contribute to individual samples, suggest that developing methods to correct for biases in sampling and normalize for the number of people contributing to a given sample will be required before these data can be acted upon in public health contexts.

@@

\begin{figure}[h]
\begin{center}
    \includegraphics[width=\textwidth]{{24hr/figures/fig5_mining_big_data.png}}
    \caption{Mining untargeted data suggests identification of human health, behavioral, and lifestyle biomarkers. Heatmap of metabolite feature abundance over 24-hour samplings, with putatively identified metabolites labeled. Metabolites labeled with filled circles were confirmed with analytical standards. Labels with open circles were identified via mass-matching neutral mass with HMDB metabolites and MS2 matching.}\label{24hr:fig5}
\end{center}
\end{figure}

### 2019-01-25 acknowledgments.tex
update 24hr with latest figures

I look back on my PhD with almost exclusive positivity: the road was rough at times, but graduate school was overall an overwhelmingly enriching and positive experience.
I recognize that this is a rare sentiment to have upon graduation, and I credit my communities, role models, mentors, family, and friends for this seemingly impossible feat.

If I have thrived in grad school, it is in large part because of my advisor, Eric Alm.
He wayers whatever seeds his students decide to plant, and I am deeply grateful for him giving me the support I needed in my research and the space I needed to succeed in my non-scientific endeavors.
Learning how to do science from Eric has given me the confidence to tackle hard problems, jump into creative solutions, and figure out the details later.
Thanks, above all, for believing in me and reminding me of that often enough for it to stick.

@@

There are many postdocs and senior graduate students whom I credit for putting me onto a path to success.
Ilana Brito was the mentor I didn't know I needed, and inspired me to hustle through this last year of my PhD to find my way to (and hopefully through) an untraditional career path.
Sarah Spencer also provided key ``unsolicited advice'' early in my rotation and thesis proposal processes which was incredibly helpful.
Her laughter is infectious and her legacy of generosity in the lab is one I aspire to.
Thomas Gurry was my first introduction to computational work during my rotation, and gave me much more time and energy than was required of him.
He provided me a solid computational foundation and was the first to introduce me to the possibility of a PhD with only computational work, which would become my passion.
In my first years in the lab, I worked closely with Mariana Matus on the Underworlds sewage project, which was key to helping me identify public health as a field where my passions and skills intersect.
Learning from her entrepreneurial spirit convinced me that there are no limits to what can be done in graduate school.
Scott Olesen was many things to me throughout my PhD: a role model, mentor, collaborator, and friend.
He was a crucially positive force in almost every aspect of my PhD: starting me on my path to student advocacy, sharing tips and tricks (and rants!) about computing, and being a model of a grad student to aspire to.
There are few people in my life who have had an impact like Scott's on shaping who I am.
Finally, I send infinite thanks to Sean Gibbons, my close collaborator on the meta-analysis and batch correction projects.
Sean is such a joy to work with, so purely believes in and loves science, and re-inspired me through my existential crises when I stopped thinking that data meant anything at all.
Beyond being just a mentor or role model, Sean is a true sponsor: volunteering me for important opportunities and always at the ready to boost my scientific career.
If more scientists (and people!) were like Sean, the world would be a better place.

@@

Beyond the work presented in these pages, I spent a significant portion of my PhD in student leadership in my department and at MIT.
Learning conflict management skills as a BE REF positively impacted almost every aspect of my life.
This impact wouldn't have happened without the people involved in my early REFS days: Libby Mahaffy, Scott, Georgia Lagoudas, Deena Rennerfeldt, Tu Nguyen, and Souparno Ghosh.
I also thank my early diversity co-conspirators, Scott Olesen and Manu Kumar, for sharing in my rants and helping me channel my passions toward productive change.
I was able to grow tremendously as a student leader because of our receptive departmental leadership, and I am very thankful to Doug Lauffenberger for being thoughtful, supportive, and responsive to our efforts.
My current GSC diversity crew, Nasir Almasri, Josue Lopez, Danielle Cosio, German Parada, Halston Lim, and Bianca Lepe, revitalized me in this work when I was closest to burning out.
The future of diversity and equity at MIT is bright if they are leaders in it.
Similarly, I am grateful for the brief time I overlapped with Jess Boles and Parker Vascik in GradSAGE, and am excited to watch them positively transform the culture of advising at MIT.

Finally, I thank my friends, family, and frisbee teammates for keeping me happy and healthy during my PhD.
I cannot thank Carolyn Saund and Megan Armstrong enough for being right by my side these past years, for supporting my growth, and for growing alongside me as we figure out how to maneuver this world.
Hold on tight to your passion, and thanks in advance for reminding me to do the same.
When I think about strong support networks of badass women, Andee Wallace and Shelby Doyle also come to mind.
Maneuvering grad school together enriched my experience tremendously, kept me focused on what matters, and was a lot of fun.
Thanks also to Brett Geiger, the BE 2014 cohort, and the entire BE community for being great friends and colleagues.
I am also thankful to Janyne Little, who keeps me adventuring and reminds me about the world outside my urban ivory towers.
Jeremy Pivor moving back to Boston was pivotal for my happiness in this city: I am so lucky for his friendship and treasure his radiant soul.
I've also been lucky to share living spaces with incredibly supportive and friendly roommates: Kunal Bhutani, a man kinder and more thoughtful than I will ever be, and the 9 Seattleites, including my dear childhood friend (and yet another badass woman inspiration) Kelly Moynihan, as well as Khoi Nguyen, Andee, Ashvin Bashyam, Griffin Clausen, Santiago Correa-Echevarria, Shawn Musgrave, and of course Nancy, Barb, Sonia, and Ruth.

To everyone I have ever thrown a frisbee with: thank you for keeping me physically and mentally healthy. I am especially grateful to have been a member of sMITe, MIT's women's ultimate frisbee team, during my entire PhD. I treasure having had the opportunity to learn about MIT's fascinating undergraduate culture and to witness the transformation of so many young women into confident players and leaders. Lisa, Bethany, Margalit, Oreo, Kimmy, Kelly, Michelle, Mary, Amy, Theresa, Sam, Ivana, and everyone else on sMITe: thanks for putting up with me, and for being my friend.

>>>>

I look back on my PhD with almost exclusive positivity: the road was rough at times, but graduate school was an overwhelmingly enriching and positive experience.
I recognize that this is a rare sentiment to have upon graduation, and I credit my communities, role models, mentors, family, and friends for this seemingly impossible feat.

If I have thrived in grad school, it is in large part because of my advisor, Eric Alm.
He waters whatever seeds his students decide to plant, and I am deeply grateful for him giving me the support I needed in my research and the space I needed to succeed in my non-scientific endeavors.
Learning how to do science from Eric has given me the confidence to tackle hard problems, jump into creative solutions, and figure out the details later.
Thanks, above all, for believing in me and reminding me of that often enough for it to stick.

@@

There are many postdocs and senior graduate students whom I credit for putting me onto a path to success.
Ilana Brito was a mentor I didn't know I needed, and inspired me to hustle through this last year of my PhD to find my way to (and hopefully through) an untraditional career path.
Sarah Spencer also provided key ``unsolicited advice'' early in my rotation and thesis proposal processes which was incredibly helpful.
Her laughter is infectious and her legacy of generosity in the lab is one I aspire to.
Thomas Gurry was my first introduction to computational work during my rotation, and gave me much more time and energy than was required of him.
He provided me a solid computational foundation and was the first to introduce me to the possibility of a PhD with only computational work, which would become my passion.
I worked closely with Mariana Matus on the Underworlds sewage project, which was key to helping me identify public health as a field where my passions and skills intersect.
Learning from her entrepreneurial spirit convinced me that there are no limits to what can be done in graduate school.
Scott Olesen was many things to me throughout my PhD: a role model, mentor, collaborator, and friend.
He was a crucially positive force in almost every aspect of my PhD: starting me on my path to student advocacy, sharing tips and tricks (and rants!) about computing, and being a model of a grad student to aspire to.
There are few people in my life who have had an impact like Scott's on shaping who I am.
Finally, I send infinite thanks to Sean Gibbons, my close collaborator on the meta-analysis and batch correction projects.
Sean is such a joy to work with, so purely believes in and loves science, and re-inspired me through my existential crises when I stopped thinking that data meant anything at all.
Beyond being just a mentor or role model, Sean is a true sponsor: volunteering me for important opportunities and always at the ready to boost my scientific career.
If more scientists (and people!) were like Sean, the world would be a better place.

@@

Beyond the work presented in these pages, I spent a significant portion of my PhD in student leadership roles in my department and at MIT.
Learning conflict management skills as a BE REF positively impacted almost every aspect of my life.
This impact wouldn't have happened without the people involved in my early REFS days: Libby Mahaffy, Scott Olesen, Georgia Lagoudas, T\'{u} Nguyen, Deena Rennerfeldt, and Souparno Ghosh.
I also thank my early diversity co-conspirators, Scott Olesen and Manu Kumar, for sharing in my rants and helping me channel my passions toward productive change.
Amanda Facklam was also key to getting the BE Application Assistance Program up and running, and I thank her for her sustained support and enthusiasm for this exciting initiative.
I was able to grow tremendously as a student leader because of our receptive departmental leadership, and I am very thankful to Doug Lauffenburger for being thoughtful, supportive, and responsive to our efforts.
My current GSC diversity crew, Nasir Almasri, Josu\'{e} Lop\'{e}z, Danielle Orozco Cos\'{i}o, German Parada, Halston Lim, and Bianca Lepe, revitalized me in this work when I was closest to burning out.
The future of diversity and equity at MIT is bright if they are leaders in it.
Similarly, I am grateful for the brief time I overlapped with Jess Boles and Parker Vascik in GradSAGE, and am excited to watch them positively transform the culture of advising at MIT.

Finally, I thank my friends, family, and frisbee teammates for keeping me happy and healthy during my PhD.
I cannot thank Carolyn Saund and Megan Armstrong enough for being right by my side these past years, for supporting my growth, and for growing alongside me as we figure out how to maneuver this world.
Hold on tight to your passion, and thanks in advance for reminding me to do the same.
When I think about strong support networks of badass women, Andee Wallace and Shelby Doyle also come to mind.
Maneuvering grad school together enriched my experience tremendously, kept me focused on what matters, and was a lot of fun.
Thanks also to Brett Geiger, the BE 2014 cohort, and the entire BE community for being great friends and colleagues.
I am also thankful to Janyne Little, who keeps me adventuring and reminds me about the world outside my urban ivory towers.
Jeremy Pivor moving back to Boston was pivotal for my happiness in this city: I am so lucky for his friendship and treasure his radiant soul.
I've also been lucky to share living spaces with incredibly supportive and friendly roommates: Kunal Bhutani, a man kinder and more thoughtful than I will ever be, and the 9 Seattleites, including my dear childhood friend (and yet another badass woman inspiration) Kelly Moynihan, as well as Khoi Nguyen, Andee Wallace, Ashvin Bashyam, Griffin Clausen, Santiago Correa-Echevarria, Shawn Musgrave, and of course Nancy, Barb, Sonia, and Ruth.

To everyone I have ever thrown a frisbee with: thank you for keeping me physically and mentally healthy. I am especially grateful to have been a member of sMITe, MIT's women's ultimate frisbee team, during my entire PhD. I treasure having had the opportunity to learn about MIT's fascinating undergraduate culture and to witness the transformation of so many young women into confident players and leaders. Lisa, Bethany, Margalit, Julia, Abby, Oreo, Kimmy, Kelly, Helen, Michelle, Mary, Amy, Theresa, Sam, Ivana, and everyone else on sMITe: thanks for putting up with me, and for being my friend.
